---
title: "Week 1 - Intro to R Syntax"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Comments

You can add comments to your code using the `#` symbol.  It can go at the
beginning of a line, or in the middle of a line after whatever code is
on that line.

```{r}
# a full line comment
1 + 1

2 + 2 # comment immediately after code
```

## The assignment operator

Any time you create a variable, or change the value of a variable, in R,
you use the arrow operator `<-` to do it.  If a line of code doesn't have
this symbol, nothing in your environment changes.

```{r}
x <- 2 + 3
```

## Indexing a single column 


You can index a column by using $ or brackets with a comma. Remember when using commas with the brackets rows are listed before the comma and columns are listed after the comma. Also when using $ you need to list the desired column name while with brackets you need to list the number associated with the column.

```{r}
#data_frame$column_name 

#data_frame[,column_number]
```

## Indexing Specific Rows and Columns


When you want to index certain rows and columns use the brackets. Numbers listed before the comman indicate the rows and numbers listed after the comma indicate columns.

```{r}
#data_frame[1:2,1:6] This will show rows 1 through 2 and columns 1 through 6
#data_frame[2,3] This will show row 2 column 3
#data_frame[2,] This will show only row 2
```

## Using c() for idexing

When you need to index certain rows or columns that aren't next to each other (so you can't used the colon) you can list your specific rows and columns using c()

```{r}
#data_frame[c(2,4,6),1:6] This will show rows 2, 4, and 6 and columns 1 through 6
```

## When you want index everything but certain items

When want index but certain items you just need to use a dash (-).
```{r}
#data_frame[-c(2,4,6),1:6] This will show columns 1 through 6 and all rows will appear except except 2, 4, and 6
#data_frame[1:2,-(1:6)] This will show rows 1 through 2 and won't show columns 1 through 6
```

## Using Brackets and Booleans to Filter Data
You can use brackets and Booleans to only show data that fit specified paratmeters.


```{r}
# CO2[CO2$Treatment == "chilled" & CO2$conc > 400,] The best example wast from lab 1. This code will return values from a dataframe that I previously named CO2 that are labelled as chilled in the Treatment column and have a value greater than 400 in the conc column.
```



