---
title: "Week 4 - working with strings"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Typing within single or double quotes will create a string 
You can also have a vector of strings
```{r echo=TRUE}
mystring <- c("B", 'and' , "C")
mystring
class(mystring)
```

To manipulate strings further we wil import data that we used in the course week 4 on compass
Take note of the option **stringsAsFactors = FALSe** this ensures character data is not converted to a categorical variable/ factor

```{r echo=TRUE}
germplasm_data <- read.table("Week_4_data/Plant.1.0 New Msa.txt",sep="\t", header = TRUE,
                             comment.char = "#", stringsAsFactors = FALSE)
```

We look at the first few lines of the data, you see we have string characters

```{r echo=TRUE}
germplasm_data[1:3,1:3]

```



String operators 
==, !=, > , <, >=, <= work


```{r echo=TRUE}
MSI <- germplasm_data$X.organism[germplasm_data$X.organism == "Miscanthus sinensis"]
head(MSI)

```

+ does not work for concatenation of strings in R
However *paste* is a handy function the can be used to concatenate strings
We will use paste to concatenate string columns in our data set

```{r echo=TRUE}
head(germplasm_data$X.sample_name)
head(germplasm_data$X.organism)
head(paste(germplasm_data$X.sample_name, ":", germplasm_data$X.organism))
head(paste(germplasm_data$X.sample_name, ", species:", 
      germplasm_data$X.organism, sep = ""))# sep"" , removes spaces between strings

```

Use paste with common separators tab(\t) and new line(\t)
```{r echo=TRUE}
new <- paste(germplasm_data$X.sample_name[1:10], collapse = "\n")
cat(new)#print line breaks

```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
