---
title: "Week 4 - working with strings"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Typing within 'single' or "double" quotes will create a string or character vector in R

You can also have a vector of strings
```{r echo=TRUE}
mystring <- c("B", 'and' , "C")
mystring
class(mystring)
```

To manipulate strings further we wil import data that we used in the course week 4.
Take note of the option *stringsAsFactors = FALSE* this ensures character data is not converted to a categorical variable

```{r echo=TRUE}
germplasm_data <- read.table("Week_4_data/Plant.1.0 New Msa.txt",sep="\t", header = TRUE,
                             comment.char = "#", stringsAsFactors = FALSE)
```

We look at the first few lines of the data, you see we have string characters

```{r echo=TRUE}
germplasm_data[1:2,1:2]

```

**String operators**

==, !=, > , <, >=, <= work

in the example below "==" is used to filter the organisms called *Miscanthus sinensis*

```{r echo=TRUE}
MSI <- germplasm_data$X.organism[germplasm_data$X.organism == "Miscanthus sinensis"]
head(MSI)

```

**paste** is a handy function the can be used to concatenate strings

"sep" option is used to add a separator between strings

```{r echo=TRUE}
paste("Miscanthus", "sinensis", sep = ":")# sep"" , removes spaces between strings
head(paste(germplasm_data$X.sample_name, germplasm_data$X.organism, sep = "_"))#use paster on a vector

```


Use paste with common separators like tab and new line
```{r echo=TRUE}
new <- paste(germplasm_data$X.sample_name[1:5], collapse = "\n")
cat(new)#print line breaks

```

*nchar* counts the number of characters in a string
it can also be used in  vectorized manner

```{r echo=TRUE}
nchar("elephantiasis")
nchar(germplasm_data$X.sample_name)#get the length of characers in a vector
```

*substring* used to subset a string, 

```{r echo=TRUE}
substring("elephantiasis", 2, 5)#2-is the start nd 5 the stop positions
head(substring(germplasm_data$X.organism, 12, nchar(germplasm_data$X.organism)))#apply to a vector

```
*strsplit* used to subset a string,

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
