---
title: "Week 4 - working with strings"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Typing within 'single' or "double" quotes will create a string or character vector in R

You can also have a vector of strings
```{r echo=TRUE}
mystring <- c("B", 'and' , "C")
mystring
class(mystring)
```

To manipulate strings further we wil import data that we used in the course week 4.
Take note of the option *stringsAsFactors = FALSE* this ensures character data is not converted to a categorical variable

```{r echo=TRUE}
germplasm_data <- read.table("Week_4_data/Plant.1.0 New Msa.txt",sep="\t", header = TRUE,
                             comment.char = "#", stringsAsFactors = FALSE)
```

We look at the first few lines of the data, you see we have string characters

```{r echo=TRUE}
germplasm_data[1:2,1:2]

```

**String operators**

==, !=, > , <, >=, <= work

in the example below "==" is used to filter the organisms called *Miscanthus sinensis*

```{r echo=TRUE}
MSI <- germplasm_data$X.organism[germplasm_data$X.organism == "Miscanthus sinensis"]
head(MSI)

```

**paste** is a handy function the can be used to concatenate strings

"sep" option is used to add a separator between strings

```{r echo=TRUE}
# sep"" removes spaces between strings
paste("Miscanthus", "sinensis", sep = ":")
#use paste on a vector to concatenate 2 columns
head(paste(germplasm_data$X.sample_name, germplasm_data$X.organism, sep = "_"))

```


Use paste with common separators like tab and new line
```{r echo=TRUE}
new <- paste(germplasm_data$X.sample_name[1:5], collapse = "\n")
#print line breaks
cat(new)

```

**nchar** counts the number of characters in a string, it can also be used in  vectorized manner

```{r echo=TRUE}
nchar("elephantiasis")
#get the length of characers in a vector
nchar(germplasm_data$X.sample_name)
```

**substring** used to subset a string

```{r echo=TRUE}
#2 and 5 are the start and stop positions
substring("elephantiasis", 2, 5)
#apply to a vector
head(substring(germplasm_data$X.organism, 12, nchar(germplasm_data$X.organism)))

```
**strsplit** splits a string into multiple strings

returns a list one vector for each original string
```{r echo=TRUE}
strsplit("this is the day", split = " ")#split by space character
strsplit("this is the day", split = " ")[[1]]

```

**formatC** and **prettyNum** used to convert numbers to text
```{r echo=TRUE}
mynum <- c(1000, 1.80, 475757686, 1.46585858589)
# scientific notation
formatC(mynum, digits = 3, format = "g")
#pad with zero on left
mynum2 <- 1:10
formatC(mynum2, flag="0", width=2)
mynum3 <- 1000:1005
prettyNum(mynum3, big.mark = ",")
prettyNum(mynum3, big.mark = " ") 

```

**match** can find the match of a value within a vector, handy for reordering data
```{r echo=TRUE}
match(c("a","b"), c("z","q","a","m","b"))

```

A vector can have **names** attributes giving names to each of its elements


```{r echo=TRUE}
myvect <- c(2,3,4,5)
names(myvect) <- c("a","b","c","d")
myvect
```


There are functions to assign or retrieve names fronm vectors: names, colnames,rownames

```{r echo=TRUE}
names(myvect)
#get column names of our dataframe
colnames(germplasm_data)
```

**Pattern Matching**

**grep** - find matched across a vector of strings, returns numeric index or the string itself
```{r echo=TRUE}
#make a vector called fruits
fruits <- c("orange", "banana", "apple")

grep("a", fruits)
grep("an", fruits)
grepl("an", fruits)
```

**find and replace** using **sub and gsub**
```{r echo=TRUE}
#substitute every match
gsub("an", "xl", fruits)
#sub first match per string
sub("an", "xl", fruits)
```

**Regular expressions**  used when you want to search for an overal pattern

can be used with grep, grepexpr, sub, gsub, strsplit to search patterns

```{r echo=TRUE}
#find a match begining with Chr
grep("^Chr", c("Chr01", "blahChr", "12Chr34")) 

#$ finds a fruit ending with e
fruits
grep("e$", fruits) 
#starts with ^a followed by any number of character then ends with $e
grep("^a.*e$", fruits)
#{2} 2 matches of an
grep("(an){2}", fruits)
```

use regular expressions to find all sample names from our dataset that stat with "JM2014-S" or "JM2014-K"

```{r echo=TRUE}
grep("^JM2014-[SK]-[0-9]{1,2}$", germplasm_data$X.sample_name,value = TRUE)
#or we can use
grep("^JM2014-[SK]-[[:digit:]]{1,2}$", germplasm_data$X.sample_name,
     value = TRUE)
```





